---
title: "Hyperion Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hyperion Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

vignette_root <- tempfile("hyperion-vignette-")
dir.create(vignette_root, recursive = TRUE)
file.copy(
  list.files(system.file("extdata", package = "hyperion"), full.names = TRUE),
  vignette_root,
  recursive = TRUE
)
file.copy(
  system.file("pharos.toml", package = "hyperion"),
  vignette_root
)
knitr::opts_knit$set(root.dir = vignette_root)
```

```{r setup}
library(hyperion)
```

# Hyperion Model object

```{r}
mod <- read_model(file.path("mod", "1001.mod"))
mod

mod <- read_model(file.path("models", "onecmt", "run002b001.mod"))
mod

mod_nm <- read_model(file.path("mod", "nmexample.mod"))
mod_nm

mod_e <- read_model(file.path("mod", "everything.mod"))
mod_e
```

```{r}
names(mod)

attributes(mod) |> names()
```

```{r}
read_model(file.path("models", "onecmt", "run001.mod")) |>
  check_dataset()

read_model(file.path("models", "onecmt", "run002.mod")) |>
  check_dataset()

read_model(file.path("models", "onecmt", "run003.mod")) |>
  check_dataset()
```


```{r, eval = FALSE}
read_model(file.path("models", "onecmt", "run001.mod")) |>
  check_model()
```
```
#> WARNINGS AND ERRORS (IF ANY) FOR PROBLEM    1
#>
#>  (WARNING  2) NM-TRAN INFERS THAT THE DATA ARE POPULATION.
#>
#> Note: Analytical 2nd Derivatives are constructed in FSUBS but are never used.
#>       You may insert $ABBR DERIV2=NO after the first $PROB to save FSUBS construction and compilation time
#> [1] 0
```

## model summary can be generated from model object

```{r}
mod <- read_model(file.path("models", "onecmt", "run003.mod"))
mod |>
	summary()
```


## parameters can be retrieved with model

```{r}
mod |> get_parameters()
```
## parameter info can be retrieved with model
```{r}
info <- get_model_parameter_info(mod)
info
```

## Model information source is captured
```{r}
mod |> get_model_parameter_info() |> audit_parameter_info()
```

# Copy model
```{r}
copy_model(
  from = file.path("models", "onecmt", "run003.mod"),
  to = file.path("models", "onecmt", "run003b2.mod"), #copies run003 to run003b1 with jittered parameters
  description = "Updating run003 to 003b1 with jittered params",
  jitter = 0.1,
  overwrite = TRUE,
  seed = 804
)
```

## Copy model accepts hyperion model object
```{r}
mod <- read_model(file.path("models", "onecmt", "run003.mod"))

mod |>
	copy_model(
		to = file.path("models", "onecmt", "run003b2.mod"),
		update = "all",
		description = "Updating run003 with mod object",
		overwrite = TRUE,
		seed = 804831
	)
```


# Managing metadata and lineage

## Set and read metadata

`set_metadata_file()` writes a JSON metadata sidecar next to a model. The
`description`, `tags`, `based_on`, and `copied_from` fields are stored
separately so model provenance can be tracked explicitly.

```{r}
set_metadata_file(
  file.path("models", "onecmt", "run003.mod"),
  description = "Base one-compartment oral absorption model",
  tags = c("base", "key"),
  based_on = c("run002.mod")
)

read_model(file.path("models", "onecmt", "run003.mod")) |>
  get_model_metadata()
```

## Populate metadata at copy time

`copy_model()` accepts `based_on` and `tags` so the new model's metadata is
populated as part of the copy.

```{r}
copy_model(
  from = file.path("models", "onecmt", "run003.mod"),
  to = file.path("models", "onecmt", "run003b2.mod"),
  description = "run003 with jittered params, exploring WT on V",
  based_on = c("run003.mod"),
  tags = c("exploratory", "wt-on-v"),
  jitter = 0.1,
  overwrite = TRUE,
  seed = 804
)
```

## Clear metadata fields

`clear_metadata_file()` selectively clears `based_on`, `copied_from`, and/or
`tags`. Fields not selected are preserved.

```{r}
clear_metadata_file(
  file.path("models", "onecmt", "run003b2.mod"),
  tags = TRUE
)
```

## Lineage queries

`get_model_lineage()` returns the project lineage tree. With no arguments
it returns every model rooted at the directory containing `pharos.toml`.

```{r}
get_model_lineage()
```

Pass a model to get its full lineage (ancestors and descendants):

```{r}
get_model_lineage(file.path("models", "onecmt", "run003.mod"))
```

Use `from` and `to` to filter the tree downward, upward, or to the slice
between two models:

```{r}
get_model_lineage(from = file.path("models", "onecmt", "run001.mod"))

get_model_lineage(to = file.path("models", "onecmt", "run003b1.mod"))

get_model_lineage(
  from = file.path("models", "onecmt", "run001.mod"),
  to   = file.path("models", "onecmt", "run003b1.mod")
)
```
